A tool set for random number generation on GPUs in R

### 2022-01-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update format
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3120 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
typo p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3122 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG. If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory. 
\fct{setBaseCreator} sets the initial state of the first stream that is created, it plays a role in \pkg{clrng} like the function \fct{SetPackageSeed} in \pkg{RngStreams} and \fct{clrngSetBaseCreatorState} in \pkg{clRNG}. \pkg{clrng} is able to create streams both on the host and on the GPU device. The \fct{createStreamsCpu} function does the former. The \proglang{R} output below creates 4 streams on the host. 
%is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior. Substreams are not created in \pkg{clrng} because currently we use one stream per work-item. 
\begin{CodeChunk}
\begin{CodeInput}
R> # creating streams on CPU
R> library("clrng")
R> setBaseCreator(rep(12345, 6))
R> myStreamsCpu <- createStreamsCpu(4)
R> t(myStreamsCpu)
\end{CodeInput}
\begin{CodeOutput}
##        [,1]       [,2]       [,3]       [,4]
##  [1,] 12345  336690377  502033783  739421137
##  [2,] 12345  597094797 1322587635 1475938232
##  [3,] 12345 1245771585 1964121530  730262207
##  [4,] 12345   85196284 1949818481 1630192198
##  [5,] 12345  523477687 1607232546  324551134
##  [6,] 12345 2094976052 1462898381  795289868
##  [7,] 12345  336690377  502033783  739421137
##  [8,] 12345  597094797 1322587635 1475938232
##  [9,] 12345 1245771585 1964121530  730262207
## [10,] 12345   85196284 1949818481 1630192198
## [11,] 12345  523477687 1607232546  324551134
## [12,] 12345 2094976052 1462898381  795289868
\end{CodeOutput} 
\end{CodeChunk}

@@

Below we generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}. The best performance we have seen using \fct{rnormGpu} is above 170 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size increases.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(512 * 128)
R> system.time(rnormGpu(c(10000, 10000), streams = streams, Nglobal = c(512,
+    128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.044   0.004   0.049
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # using GPU
R> streams <- createStreamsGpu(n = 256*64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- 
+  clrng::fisher.sim(month_gpu, 1e6, 
+     streams=streams, type="double", 
+     returnStatistics=TRUE,  Nglobal = c(256,64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.335   0.014   0.348
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # using GPU
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(resultWeek <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.934   0.014   1.947
\end{CodeOutput}
\begin{CodeInput}
R> unlist(resultWeek[c("threshold", "simNum", "counts")])
\end{CodeInput}
\begin{CodeOutput}
## threshold    simNum    counts 
##    -54990  10010624      1350
\end{CodeOutput}
\begin{CodeInput}
R> resultWeek$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{figure}[t]

>>>>

%The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG. If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory. 
\fct{setBaseCreator} sets the initial state of the first stream that is created, it plays a role in \pkg{clrng} like the function \fct{SetPackageSeed} in \pkg{RngStreams} and \fct{clrngSetBaseCreatorState} in \pkg{clRNG}. \pkg{clrng} is able to create streams both on the host and on the GPU device. The \fct{createStreamsCpu} function does the former. The \proglang{R} output below creates 4 streams on the host. 
%is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior. Substreams are not created in \pkg{clrng} because currently we use one stream per work-item. 
\begin{CodeChunk}
\begin{CodeInput}
R> # creating streams on CPU
R> library("clrng")
R> setBaseCreator(rep(12345, 6))
R> myStreamsCpu <- createStreamsCpu(4)
R> t(myStreamsCpu)
\end{CodeInput}
\begin{CodeOutput}
##               [,1]       [,2]       [,3]       [,4]
## current.g1.1 12345  336690377  502033783  739421137
## current.g1.2 12345  597094797 1322587635 1475938232
## current.g1.3 12345 1245771585 1964121530  730262207
## current.g2.1 12345   85196284 1949818481 1630192198
## current.g2.2 12345  523477687 1607232546  324551134
## current.g2.3 12345 2094976052 1462898381  795289868
## initial.g1.1 12345  336690377  502033783  739421137
## initial.g1.2 12345  597094797 1322587635 1475938232
## initial.g1.3 12345 1245771585 1964121530  730262207
## initial.g2.1 12345   85196284 1949818481 1630192198
## initial.g2.2 12345  523477687 1607232546  324551134
## initial.g2.3 12345 2094976052 1462898381  795289868
\end{CodeOutput} 
\end{CodeChunk}

@@

Below we generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}. The best performance we have seen using \fct{rnormGpu} is above 170 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size increases.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(512 * 128)
R> system.time(rnormGpu(c(10000, 10000), streams = streams, Nglobal = c(512,
+    128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.044   0.005   0.049
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # using GPU
R> streams <- createStreamsGpu(n = 256*64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- 
+  clrng::fisher.sim(month_gpu, 1e6, 
+     streams=streams, type="double", 
+     returnStatistics=TRUE,  Nglobal = c(256,64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.317   0.000   0.318
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # using GPU
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(resultWeek <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.942   0.015   1.955
\end{CodeOutput}
\begin{CodeInput}
R> unlist(resultWeek[c("threshold", "simNum", "counts")])
\end{CodeInput}
\begin{CodeOutput}
## threshold    simNum    counts 
##    -54990  10010624      1350
\end{CodeOutput}
\begin{CodeInput}
R> resultWeek$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{figure}[h]
